---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# easydb

DBI and dplyr wrappers to write to DB, fetch and run data manipulation operations on server side.

We define `$.DBIConnection()`, `$<-.DBIConnection()` and
`!.tbl_lazy()` to be able to :
* write in a data base with
  `con$foo <- my_data_frame` (where `con` is a connection built with
  `DBI::dbConnect()` or this package's db_connect), short for
  `DBI::dbWriteTable(con, "foo", my_data_frame)`
* build a lazy table with `con$foo`, short for `dplyr::tbl(con, "foo")`,
  so it can be used to do some server side data manipulation using *dplyr*
* collect a lazy table into the R session with `!my_lazy_table` (or directly
  `!con$foo`), equivalent to `dplyr::collect(my_lazy_table)` or
  `dplyr::collect(dplyr::tbl(con, "foo"))`

## Installation

Install with :

    remotes::install_github("moodymudskipper/easydb")

## connect

We offer a simple wrapper around `DBI::dbConnect` that makes it easy to use 
specs stored in a list and makes it easier to choose the relevant driver.

```{r}
library(easydb)
# Used just as DBI::dbConnect()
db <- db_connect(RSQLite::SQLite(), path = ":memory:")
# Using shorthand
db <- db_connect("sqlite", path = ":memory:")
# Using a list of specs
specs <- list(drv = "sqlite", path = ":memory:")
db <- db_connect(specs)
```

## Query

To write or fetch :

```{r}

db$foo <- head(cars,2)

db$foo

!db$foo


```
