---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
requireNamespace("dbplyr")
```
# easydb

DBI and dplyr wrappers to write to DB, fetch and run data manipulation operations on server side.

We define `$.DBIConnection()`, `$<-.DBIConnection()` and
`!.tbl_lazy()` to be able to :

* write in a data base with
  `con$foo <- my_data_frame` (where `con` is a connection built with
  `DBI::dbConnect()` or this package's db_connect), short for
  `DBI::dbWriteTable(con, "foo", my_data_frame)`
* build a lazy table with `con$foo`, short for `dplyr::tbl(con, "foo")`,
  so it can be used to do some server side data manipulation using *dplyr*
* collect a lazy table into the R session with `!my_lazy_table` (or directly
  `!con$foo`), equivalent to `dplyr::collect(my_lazy_table)` or
  `dplyr::collect(dplyr::tbl(con, "foo"))`
* We can access tables in a schema by using a `.` separator as we would in
  an SQL query. In that case `con$my_schema.my_table` is equivalent to
  `dplyr::tbl(con, dbplyr::in_schema("my_schema", "my_table"))`

## Installation

Install with :

    remotes::install_github("moodymudskipper/easydb")

## connect

We offer a simple wrapper around `DBI::dbConnect` that makes it easy to use specs stored in a list and makes it easier to choose the relevant driver.

```{r}
library(easydb)
# Used just as DBI::dbConnect()
db <- db_connect(RSQLite::SQLite(), path = ":memory:")
# Using shorthand
db <- db_connect("sqlite", path = ":memory:")
# Using a list of specs
specs <- list(drv = "sqlite", path = ":memory:")
db <- db_connect(specs)
# Or just, for this specific case, just use the defaults and it will create
# an SQLite database in memory
db <- db_connect()
```

## Query

To write or fetch :

```{r}
db$foo <- head(cars,2)
db$bar <- head(iris,2)
dplyr::db_list_tables(db)
db$foo

!db$foo
```

By default one can't overwrite a table, it would need to be removed first, which
you can do by calling `db$table_to_delete <- NULL`

```{r, error=TRUE}
db$foo <- tail(cars,2)
db$foo <- NULL
db$foo <- tail(cars,2)
!db$foo
```

We can allow overwriting by setting the option `easydb.overwrite` to `TRUE`.

```{r}
options(easydb.overwrite = TRUE)
db$bar <- tail(iris,2)
!db$bar
```

We can attach or create a schema with the following syntax (here creating a
schema named `"aux"` in a temporary file) :

```{r}
db$aux. <- tempfile()
```

Then we can use a smilar syntax to what's already been showed, by prefixing
table names with the schema name and a dot.

```{r}
db$aux.baz <- head(mtcars[1:4], 2)
!db$aux.baz
```

## using `with()`

We provide a method `with.DBIConnection()` that can be used to do operators "in"
the database. Just as `base:::with.default()` looks for the symbol of the expression in its first argument first, `with.DBIConnection()` searches for the tables first in the database using the standard `with()`.

A difference is that one can assign inside of the database directly by using `<-` in the expression.

```{r}
with(db,{
   # new table from r data
  baz <- head(mtcars,2)
  # create new table from dbms data, all computed on server side
  qux <- dplyr::union_all(foo,foo) 
})
db$baz
db$qux
```


